\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-14 k-planar.tex
Pfffft

\begin{thm}
\label{kPlanarProduct}
Every $k$-planar graph is contained in $H\boxtimes P\boxtimes K_{18k^2+48k+30}$, for some graph $H$ of treewidth $\binom{k+4}{3}-1$ and for some path $P$.
\note{DW}{Do we want to mention ``every $1$-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ of treewidth $3$ and for some path $P$'' here?}
\end{thm}

@@

For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and whose crossing edges are in the interior of these faces \note{DW}{I suggest we replace ``and whose crossing edges are in the interior of these faces'' by ``such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$. } The graph $G_0$ is called the \defin{frame} of $G_0$. This definition (for simple graphs) was introduced by \citet{BDGGMR}. Every $d$-framed multigraph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar graph. Thus \cref{ShortcutProduct} is applicable. However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible\footnote{A multigraph $G$ is \defin{contained} in a graph $G'$ if the simple graph underlying $G$ is isomorphic to a subgraph of $G'$.}.

@@

\begin{lem}
	\label{induction} The setup:
	\begin{compactenum}
		\item Let $G$ be a $d$-framed graph and let $G'$ be the frame of $G$.
		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that $V(F)$ can be partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
		\begin{compactenum}
			\item $F[P_i]$ is a path; and
			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
		\end{compactenum}
		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
	\end{compactenum}
	Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
	\begin{compactenum}[(i)]
		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
		\item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
	\end{compactenum}
\end{lem}

@@

@@

>>>>

\begin{thm}
\label{kPlanarProduct}
Every $k$-planar graph is contained in $H\boxtimes P\boxtimes K_{18k^2+48k+30}$, for some graph $H$ of treewidth $\binom{k+4}{3}-1$ and for some path $P$.
\end{thm}

\note{DW}{Do we want to mention ``every $1$-planar graph is contained in $H\boxtimes P\boxtimes K_7$ for some planar graph $H$ of treewidth $3$ and for some path $P$'' here?}

\note{PM}{Yes, I think so.  Two theorems, like this, or one?}

For the important special case of $1$-planar graphs, we are able to obtain better constants and an additional property (planarity) on $H$.

\begin{thm}\label{1_planar_product}
  Every $1$-planar graph is contained in $H\boxtimes P\boxtimes K_{7}$, for some planar graph $H$ of treewidth $3$ and for some path $P$.
\end{thm}

@@

For any integer $d\ge 3$, a (multi)graph $G$ is \defin{$d$-framed} if it admits a drawing in the plane having a biconnected plane (multi) subgraph $G_0$, whose facial cycles have length at most $d$,  and such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$.
% \note{DW}{I suggest we replace ``and whose crossing edges are in the interior of these faces'' by ``such that each edge in $E(G)\setminus E(G_0)$ is in the interior of some face of $G_0$. }
The graph $G_0$ is called the \defin{frame} of $G_0$. This definition (for simple graphs) was introduced by \citet{BDGGMR}. Every $d$-framed multigraph can be described by a $(2,\binom{d}{2})$-shortcut system applied to a planar graph. Thus \cref{ShortcutProduct} is applicable. However, for framed graphs we prove the following stronger result, where the treewidth bound on $H$ is best possible\footnote{A multigraph $G$ is \defin{contained} in a graph $G'$ if the simple graph underlying $G$ is isomorphic to a subgraph of $G'$.}.

@@

\begin{lem}
	\label{induction} The setup:
	\begin{compactenum}
		\item Let $G$ be a $d$-framed multigraph and let $G'$ be a frame of $G$.
		\item Let $T$ be a BFS spanning tree of $G'$ rooted at some vertex $r$.
		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(r,v)=j\}$.
		\item Let $F$ be a cycle in $G'$ with $r$ in the exterior of $F$ and such that $V(F)$ can be partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
		\begin{compactenum}
			\item $F[P_i]$ is a path; and
			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
		\end{compactenum}
		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
	\end{compactenum}
	Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
	\begin{compactenum}[(i)]
		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
		\item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
	\end{compactenum}
\end{lem}

@@

% \todo[inline]{Alternative to the final three paragraphs in the preceding proof:
%
% To see that $H$ is planar, consider the plane multigraph $G''$ obtained by adding a vertex in the interior of each face of $G'$.
% }

@@

\note{PM}{Annoying detail:  When we allow framed graphs to have crossings in the outer face, we can't guarantee that $G'$ has any face smaller than $d$. This makes it awkward to start the induction using \cref{induction}. Possible solutions:
\begin{compactenum}
  \item Disallow the addition of edges in the outer face of a frame.
  \item Modify \cref{induction} so that $T$ is a BFS forest rooted at the set of vertices on the outer face of $G'$. Then we can start the induction with $k=1$.
\end{compactenum}
I prefer the second option.  Thoughts?
}
